---
title: "Getting start"
author: "Pablo Garcia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignette we provide a short of description on how \code{firstR} works.

The package allows to summarize and visualize data from the [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

# Data loading

Some sample data is included in the package for convenience. It includes fars data from years 2013 to 2015.

Data is stored in *.csc.bz2 files in the `r file.path('inst','extdata')` directory, to get start copy any of the files to your working directory.

This files have a standard name, so you you can access data from the package functions using only the year number as input.

# Reading full datasets

To get the full data set use the function \code{\link{fars_data_years}}, this 
function returns a list of tibble, one for each year of data.

```{r}
library(firstR)
fars_2014 <- fars_read_years(years = c(2014,2015))

head(fars_2014)
```

# Summarizing data
 
A summary of the data in terms of number of fatalities by month can be computed with \code{\link{fars_summarize_years}}

The following code computes the summary for 2013 data.

```{r}

fars2013_14 <- fars_summarize_years(years = c(2013,2014))

fars2013_14

```

# Data visualization

Finally the accident location can be plotted for a single state input. State code must be used as input.

```{r}
fars_map_state(state.num = 1, year = 2013)

```
